Pool Model Paper

### 2023-12-06 pool_model.tex
Correct some small typos

\section{One Species Modelling}

@@

When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurres.

@@

To detremine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.

@@

For now we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 

@@

Note that the pools $L$, $G$ and $D$ represent number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is number of bacteria and $V$ is the volume. 

@@

By defeinition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 

@@

One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
Moreover, the rapid shift in temperature is associated with the appearence of an additional lag phase~\cite{zwietering_modeling_1994}.

@@

        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).

@@

        The paremeters of the back-flow rate are $\Gamma=1$ and $\delta=5$.

@@

Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species. 

@@

One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
To descibe such a situation we consider for simplicity only one common limiting resource pool. 

@@

which has the structure of a generalised Lotka-Volterra system (gLV).

@@

Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
In previous studies it was shown that gLV equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
In food studies the gLV was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.

@@

\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}

@@

The function $\mathcal{B}_{A/B}$ captures the inhibiting effects in an effective way. 

@@

On the other hand, one can also observe some cooperative behaviours between species in case when it facilitates the competition with others.

@@

Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself \cite{rainey_evolution_2003}.

@@

- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modelling (ABM).\\
More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.

@@

    \item \cite{rosser_modelling_2014}: modelling of the bacterial reorintation, rotation.
    \item \cite{sokolov_physical_2012, li_amplified_2008}: modelling of the bacteria using hydrodynamic interactions and Brownian motion.

@@

    \item \cite{farrell_mechanically_2013}: 2D modelling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.

@@

    The growth of bacterial colonies where nonmotile microorganisms replicate and push each other away as they grow.

>>>>

\section{One Species Modeling}

@@

When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurs.

@@

To determine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.

@@

For now, we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 

@@

Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume. 

@@

By definition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 

@@

One of the reasons of the dormant state appearance can be also the temperature shift~\cite{oliver_viable_1995}.
Moreover, the rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.

@@

        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ODE_tempshift}).

@@

        The parameters of the back-flow rate are $\Gamma=1$ and $\delta=5$.

@@

Further we show that the TPM can be readily extended to the systems with more than one bacteria and describe more complex relationships between species. 

@@

One on the ways to describe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
To describe such a situation we consider for simplicity only one common limiting resource pool. 

@@

which has the structure of a \ac{glv}.

@@

Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
In previous studies it was shown that \ac{glv} equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.

@@

\cmt{\cite{lenski_coexistence_1986}: Modeling of the 2 sp. system with 1resource and 1 inhibitor.}

@@

The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively. 

@@

On the other hand, one can also observe some cooperative behaviors between species in case when it facilitates the competition with others.

@@

Sometimes neighbors, so-called 'social cheaters', can benefit from a certain behavior without expressing it by itself \cite{rainey_evolution_2003}.

@@

- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.

@@

    \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
    \item \cite{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.

@@

    \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.

@@

    The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
